---
title: "Appendix of requested variables"
author:
  - name: Luke W. Johnston
    url: https://lukewjohnston.com/
    affiliation: Steno Diabetes Center Aarhus
repository_url: https://gitlab.com/lwjohnst/meld-dst-application/
creative_commons: CC BY
---

`r if (!knitr::is_html_output()) " TODO: Delete above title header before sending to DST. "`

Danmarks Statistik, Forskningsservice
`r lubridate::stamp("March 13, 1999", quiet = TRUE)(lubridate::today())`\
Generel Forskningsservice\
Projekt nr. 708421

Luke W. Johnston, Steno Diabetes Center Aarhus

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(glue)
library(flextable)
source(here("R/format-tables.R"))
needed_registers <- here("data/needed-registers.csv") %>% 
    read_csv(col_types = "lddccccc") %>% 
    filter(!drop | is.na(drop))

dst_variables <- read_csv(here("data/dst-registers-with-variables-to-use.csv"))
dst_variables <- dst_variables %>%
    filter(use_in_application == 1,
           register_id %in% unique(needed_registers$register_id)) %>% 
    mutate(across(ends_with("year"), str_extract, pattern = "^\\d{4}")) %>% 
    mutate(
        start_year = case_when(
            is.na(variable_start_year) |
                variable_start_year == variable_end_year ~
                register_start_year,
            TRUE ~ variable_start_year
        ),
        year_range = glue::glue("{start_year} - {variable_end_year}")
    ) %>% 
    select(register_id, register_name_dk, variable_name, year_range, description_dk)

sds_variables <- read_csv(here("data/sds-registers-with-variables-to-use.csv"))
sds_variables <- sds_variables %>% 
    filter(use_in_application == 1) %>% 
    # mutate(start_year = str_extract(periode, "^\\d{4}")) %>% 
    mutate(periode = str_replace(periode, "\\s?-\\s?$", "-Latest")) %>% 
    select(register_name, table_name, variable_name, periode, variablebeskrivelse)

random_registers <- read_csv(here("data/random-undocumented.csv")) %>% 
    mutate(
        start_year = if_else(is.na(start_year), "Earliest", as.character(start_year)),
        year_range = glue::glue("{start_year} - {end_year}")
    )

knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")
# To get loop to work
if (knitr::is_html_output())
    htmltools::div(flextable_html_dependency())
```

## Projekttitel

Interplay between diabetes and intergenerational transmission of health
determinants over the life course

## Populations afgrænsning(er)

<!-- Skriv her en beskrivelse af populationen -->

In create the family structure and linkages, we need:

-   All index individuals born or who migrated into Denmark on or after
    1970-01-01.
-   Registers and variables:
    -   BEF: AEGTE_ID, CIVST, CIV_VFRA, E_FAELLE_ID, FAR_ID, MOR_ID,
        PNR, FOED_DAG, KOEN, OPR_LAND.
    -   CPST: AEGTEFAELLE_ID, CIVST, CIV_DATO, FAR_ID, MOR_ID, PNR.

Combining BEF and CPST has, from our previous experience and projects,
resulted in better coverage in creating our study population than one
register alone.

## List of registers and variables from DST

```{r requested-dst-variable-tables, results='asis', echo=FALSE}
variable_tables <- dst_variables %>% 
    mutate(register_name = glue("{register_name_dk} ({register_id})")) %>% 
    select(-register_name_dk) %>% 
    rename(Variables = variable_name, Beskrivelse = description_dk, År = year_range) %>% 
    group_split(register_name) 

for (variable_table in variable_tables) {
    register <- unique(variable_table$register_name)
    # Register name as header
    glue("\n\n## Grundregisternavn: {register}\n\n") %>% 
        as.character() %>% 
        cat()
    
    # Variables as table
    variable_table %>% 
        format_dst_appendix_table()
}
```

## List of registers and variables from external sources

```{r requested-non-dst-variable-tables, results='asis', echo=FALSE}
sds_variable_tables <- sds_variables %>% 
    rename(
        Tabel = table_name,
        Variabler = variable_name, 
        Beskrivelse = variablebeskrivelse,
        År = periode
    ) %>% 
    filter(register_name == "Det Nationale Diabetesregister")

register <- unique(sds_variable_tables$register_name)
# Register name as header
glue("\n\n## {register}\n\n") %>%
    as.character() %>%
    cat()

# Variables as table
sds_variable_tables %>%
    dplyr::select(matches("^Tabel$"), Variabler, Beskrivelse, År) %>% 
    format_non_dst_appendix_table(width = c(1.5, 1.5, 3, 1.5))
```

```{r requested-random-registers-variable-tables, results='asis', echo=FALSE}
random_variable_tables <- random_registers %>% 
    select(register_name, register_table_name, variable_name, 
           year_range, variable_description_dk) %>% 
    rename(
        Tabel = register_table_name,
        Variabler = variable_name, 
        Beskrivelse = variable_description_dk,
        År = year_range
    ) %>% 
    group_split(register_name) 

for (variable_table in random_variable_tables) {
    register <- unique(variable_table$register_name)
    # Register name as header
    glue("\n\n## {register}\n\n") %>% 
        as.character() %>% 
        cat()
    
    width <- c(1.5, 1.5, 3, 1.5)
    if (all(is.na(unique(variable_table$Tabel)))) {
        width <- 2.25
        variable_table <- variable_table %>% 
            select(-Tabel)
    }
    
    # Variables as table
    variable_table %>% 
        dplyr::select(matches("^Tabel$"), Variabler, Beskrivelse, År) %>% 
        format_non_dst_appendix_table(width = width) 
}
```

## Labka

For alle individer i databasen ønskes der laboratorie udtræk af følgene
NPU/DNK-koder så langt tilbage i tiden som muligt.

```{r labka-data-source, results='asis'}
here("data/labka.csv") %>%
    read_csv(col_types = cols(
        Komponent = col_character(),
        Kode = col_character(),
        Enhed = col_character()
    )) %>% 
    group_by(Komponent) %>% 
    mutate(Komponent = if_else(row_number() == 1, Komponent, "")) %>% 
    format_non_dst_appendix_table()
    # knitr::kable(caption = "Components and codes from the Labka database.")
```

## List of requested ICD-8 and -10 codes

Da hovedformålet med projektet er at undersøge prævalensen og incidensen
af diabetes, relaterede kardiometaboliske sygdomme samt følgesygdomme
(komplikationer) i sociale netværk ønsker vi nedenstående ICD10 samt
ICD8 koder. Efter koderne har vi kort redegjort for, hvorfor vi ansøger
om netop de listede koder.

```{r icd-codes-tables, results='asis'}
icd_codes <- here("data/icd-codes.csv") %>%
    read_csv(
        col_types = cols(
            parent_group = col_character(),
            Sygdomsgruppe = col_character(),
            Sygdomme = col_character(),
            ICD10 = col_character(),
            ICD8 = col_character()
        )
    )

icd_as_table(
    icd_codes,
    "diabetes",
    "Diagnosis of diabetes and its forms and subtypes."
)

icd_as_table(
    icd_codes,
    "comorbidity",
    "Diseases that contribute to, are associated with, or are complications or comorbidities of diabetes."
)

icd_as_table(
    icd_codes %>% select(-Sygdomsgruppe),
    "charlesons",
    "Diseases that are components of the 'Charleson's Morbidity Index'.",
    width = 2.25
)
```

## Justification for requested disease codes

The overall aim of this project is to identify the contributions of
family and early life determinants on the development, management and
care of diabetes and the diseases that may arise following a diabetes
diagnosis, under a life-course framework. The justification for
requesting data on multiple disease areas directly follows the primary
aim of this project. We investigate this aim focusing on diabetes as the
central disease of interest. Diabetes consists of different subtypes
(Type 1, Type 2, LADA, MODY, gestational diabetes, secondary diabetes,
rare monogenic forms) each with a different set of risk factors,
presentation and pathophysiological characteristics. The familial and
social effects that are the subject of this project are likely to be
different for each of these diabetes subtypes. E.g. for some types of
diabetes, caused predominantly by auto-immune mechanisms, familial
associations to other auto-immune diseases are more likely than for type
2 diabetes, which is driven to a strong degree by obesity, low physical
activity and insulin resistance.

All forms of diabetes have an increased blood glucose level as their
central hallmark, and are associated with major and minor complications.
The vascular complications are generally subdivided into large-vessel
disease (macrovascular): myocardial infarction, stroke, peripheral
vascular disease; and small-vessel disease (diabetic retinopathy,
neuropathy nephropathy). The occurrence of these complications is not
dependent solely on the elevated glucose levels, but also on disturbance
of other metabolic risk factors, familial predisposition and
pre-existing conditions. Beyond these classical complications,
increasingly links between diabetes and other complications are being
recognised: depression, cancer, loss of cognitive function, skin
conditions.

An important feature of diabetes is that it can be undetected for
several years, and that the diabetic complications can sometimes be the
first presentation of the disease. In order to study the occurrence of
diabetic complications in the context of family we need to assess the
complication status for all traditional and novel diabetic complication.

Adequate treatment of diabetes depends on long-term engagement and
motivation of the patient for self-management of different aspects of
the disease. The capacity to respond adequately to this challenge
depends to a large degree on socio-economic status, including the degree
of social support in the direct environment surrounding the patient.
These effects occur in interaction with other chronic conditions,
including mental health conditions.

The justification for the requested list of conditions falls into four
categories:

1.  Diagnosis of diabetes itself, in all its forms and subtypes.
2.  Conditions that are an established cause of diabetes, and diseases
    which have an emerging association with diabetes, which we wish to
    investigate
3.  Major and minor complications and other consequences / signs of
    diabetes
4.  Conditions needed to adjust our analyses for the simultaneous
    occurrence of other chronic health problems (co-morbidity)

Description of disease and its relation to diabetes

-   **Cardiovascular disease** is the most common cause of death and
    disability among people with diabetes The cardiovascular diseases
    that accompany diabetes include angina, myocardial infarction (heart
    attack), stroke, peripheral artery disease and congestive heart
    failure. High blood pressure, high cholesterol, high blood glucose
    and other risk factors contribute to the increased risk of
    cardiovascular complications.

-   **Eye diseases:** Diabetic retinopathy and macular edema are the
    major eye complications of diabetes; but diabetic eye disease also
    includes cataract and glaucoma. Furthermore, infections of the
    eyelid and adnexa are also seen.

-   **Neurological and neurodegenerative diseases:** Diabetic peripheral
    and autonomic neuropathy are part of the major diabetic
    microvascular complications. Furthermore, diabetes has long been
    linked to vascular dementia; probably mediated through chronic
    ischaemia, endothelial dysfunction and micro-strokes. In the past
    decade, increasing evidence has emerged for a shared
    pathophysiological and aetiological connection between diabetes and
    Parkinson's disease, Alzheimer's disease, dementia, loss of
    cognitive function.

-   **Mental health conditions** have been linked to diabetes in
    different ways. The most established link is between depression and
    diabetes, both as a cause and a consequence, but also an association
    with schizophrenia has been observed and is receiving increasing
    attention. Mental health can deteriorate as a consequence of living
    with diabetes, and poor mental health, including personality and
    eating disorders can impair a patient's ability to cope adequately
    with the demands (changes in lifestyle, self-measurement and
    management, self-care) posed by the presence of long-standing
    diabetes. Consequently, patients with co-morbid diabetes and mental
    health problems tend to have worse control and outcomes for both
    conditions. Moreover, mental health conditions also affect people's
    abilities to build up and maintain social connections, and they may
    place strain on family relations, potentially altering their
    influence on health compared to the general population.

-   **Orodental health:** There is a well-established link between
    diabetes and orodental health. Diabetes is a risk factor for
    periodontal diseases such as gingivitis and periodontitis, and the
    relation is though to be bidirectional, meaning that not only does
    diabetes increase the occurrence of orodental problems, but also
    that orodental conditions may affect the metabolic control of
    diabetes in diabetic patients.

-   **Fractures** are associated with diabetes both directly and
    indirectly (as a consequence of loss of eyesight or loss of
    sensation in the feet due to diabetic retinopathy and neuropathy).
    Moreover, fractures, especially in the elderly are a strong cause of
    co-morbidity and can exacerbate other chronic health problems.

-   **Gastro-intestinal diseases:**

    -   Diabetes is ultimately a disease of imbalance in the
        carbohydrate metabolism, and is closely linked to lipid
        metabolism. Besides the liver, there is an increasing insight
        that the gut plays an important regulatory role. Several
        mechanisms linked to appetite and satiety have been found to be
        regulated in feed-back loops involving the gut (particularly the
        duodenum). This is expressed in the marked improvement in
        metabolic control seen in patients with diabetes who undergo
        Roux-en-Y gastric bypass operations and the impact on obesity of
        a Glucagon-like peptide-1 analogues.
    -   Another emerging line of research linking the health of the gut
        to diabetes is the study of the intestinal microbiome. This
        emerging field is showing increasingly convincing associations
        between diet, the composition of the bacterial populations in
        the gut and obesity, CVD and several metabolic diseases
        including diabetes. It follows that any condition that causes
        gut inflammation, and malabsorption, such as Irritable Bower
        Syndrome, Crohn's disease or gastric and duodenal ulcers, may
        affect the intestinal microflora, and through that mechanism
        diabetes risk.
    -   Existing diabetes can also cause gut symptoms; patients with
        severely disregulated diabetes sometimes present with
        gastroparesis and constipation, which can be seen as a
        consequence of failed autonomic regulation of gut motility due
        to diabetic autonomic neuropathy.

-   **Liver diseases:** Insulin resistance, a central determinant of
    type 2 diabetes, is principally determined by insulin resistance in
    the muscles (peripheral) and in the liver (central). Diseases that
    affect liver function, through inflammation, fat accumulation,
    fibrosis or a combination of these mechanisms, have an impact of
    central insulin resistance. This phenomenon is recognised in the
    strong relationship between Non-Alcoholic Fatty Liver Disease
    (NAFLD) and Non-Alcoholic Steatohepatitis (NASH) with diabetes and
    the metabolic syndrome. The central role of the liver, and liver
    conditions, to diabetes, also provides a conceptual link to other
    metabolic conditions in which the liver plays a central role, such
    as iron and lipid metabolism.

-   **Kidney diseases:** Diabetic kidney disease is one of the main
    microvascular complications of diabetes, progressing from a mild
    initial presentation (micro-albuminuria) to advanced
    (macro-albuminuria) and late stages (end-stage renal disease,
    ultimately with renal failure and the need for dialysis). Loss of
    renal function itself (both in the presence and absence of diabetes)
    is recognised as a risk factor for the development of cardiovascular
    disease. Diabetic kidney disease thus also provides a direct
    pathophysiological link to the management of hypertension in people
    with diabetes.

-   **Anaemia and Iron metabolism:** Iron metabolism has long been
    implicated as a causal factor in cardiovascular disease aetiology,
    but there is mounting evidence pointing to a causal link the
    occurrence of diabetes. Low iron stores and iron deficiency anaemia
    can have an impact on the ability to use HbA1c as a valid indicator
    of average glucose levels over a longer period of time. Anaemia is
    also a strong indicator of general ill health and should be
    considered when studying diabetes in a context of multi-morbidity.

-   **Skin diseases:** Several dermatological conditions, including
    psoriasis, dermatitis and eczema, are seen more frequently in
    patients with diabetes. Peripheral neuropathy affects the skin's
    ability to activate sweat glands, leading to dry and brittle skin
    and higher occurrence of skin infections, which exacerbate diabetic
    foot problems.

-   **Infertility:** Fertility problems are well known for women with
    diabetes and male infertility is a likely but less studied
    consequence of diabetes.

## List of requested ATC codes

```{r atc-code-table, results='asis'}
here("data/atc-codes.csv") %>%
    read_csv(col_types = cols(
        level = col_double(),
        code = col_character(),
        description = col_character()
    )) %>% 
    select(`ATC Code` = code, Description = description) %>% 
    format_non_dst_appendix_table(caption = "Requested ATC codes in classifying diabetes and cardiovascular diseases.", width = c(2, 5), is_atc = TRUE)
```
