---
title: "Project description for DST application"
author:
  - name: Luke W. Johnston
    url: https://lukewjohnston.com/
    affiliation: Steno Diabetes Center Aarhus
repository_url: https://gitlab.com/lwjohnst/meld-dst-application/
creative_commons: CC BY
---

`r if (!knitr::is_html_output()) "
TODO: Delete above title header before sending to DST.
"`

**Forskningsservice**                     **Dato** `r lubridate::stamp("March 13, 1999", quiet = TRUE)(lubridate::today())`
Projekt nr. 708421/ kontaktperson i Forskningsservice

`r if (!knitr::is_html_output()) "
TODO: Fix kontaktperson line.
"`

## Autoriseret institution / Authorised institution

Enhedschef Annelli Sandbæk  
Steno Diabetes Center Aarhus (inst. nr. 574)  
Hedeager 3, 2. sal  
8200 Aarhus  
Denmark  

## Projekttitel / Project title

Interplay between diabetes and intergenerational transmission of health determinants over the life course

## Projektbeskrivelse / Project description

The overall aim of this project is to identify the contributions of family and
early life determinants on the development, management, and care of diabetes and
the diseases that may arise following a diabetes diagnosis, under a life-course
framework.

Diabetes is a complex, multi-factorial disease. Its development depends on the
timing and accumulation of exposures throughout the life-course. In addition to
genetic influences, behavioral risk factors and socioeconomic conditions cluster
in families and are passed through generations. For instance, poor childhood
socioeconomic status (SES) can impact final educational attainment, adult SES,
and lifestyle and behavioral characteristics, which ultimately influence an
individual's metabolic capacity to regulate glucose and subsequent risk for
diabetes.

This project will contribute to a broader and deeper understanding of the role
that these factors play with the development and progression of diabetes. For
this purpose, we will apply causal and predictive statistical methods to best
use the data available from the integrated data sources with an overarching goal
to model several potential public health strategies that might impact the
prevention and management of diabetes and the diseases that follow its
diagnosis. 

<!-- A key feature to this project is the integration of external data -->
<!-- sources, such as longitudinal cohort studies, with the register data. -->

## Population / Population

In order to identify familial relations through individuals' life course trajectory,
the population requested covers all individuals, who were alive in Denmark 
from 1 January 1970 to 31 December 2020 (~8.5 million people). 
The family linkage and structure makes the construction of the population
complicated and we will do this ourselves. For that we need the entire
population because we need data on not just the individual, but also the family
(parents and siblings). To characterise early life and early childhood
conditions of an individual person, we need data on their parents. Likewise, to
understand the role that the family network, including siblings and
grandparents, has on diabetes risk, we need a large enough population to make
these linkages and to adequately study this area.
Key requirements are:

- All index individuals born or who migrated into Denmark.
- All family members of the index individuals, including parents, siblings,
spouses, and children.

## Variabelindhold / Variables included

All variables can be de-identified (direct identification is not needed). The
list of requested variables is found in the tables at the end of the document in 
the Appendix. For all the variables, we need the data from the first year it is
available to the most recent year it is available. 

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(glue)
needed_registers <- here("data/needed-registers.csv") %>% 
    read_csv(col_types = "lddccccc") %>% 
    filter(!drop | is.na(drop))

dst_variables <- read_csv(here("data/dst-registers-with-variables-to-use.csv"))
dst_variables <- dst_variables %>%
    filter(use_in_application == 1,
           register_id %in% unique(needed_registers$register_id)) %>% 
    mutate(across(ends_with("year"), str_extract, pattern = "^\\d{4}")) %>% 
    mutate(start_year = case_when(
        is.na(variable_start_year) | variable_start_year == variable_end_year ~
            register_start_year,
        TRUE ~ variable_start_year
    )) %>% 
    select(register_id, register_name_dk, variable_name, start_year)

sds_variables <- read_csv(here("data/sds-registers-with-variables-to-use.csv"))
sds_variables <- sds_variables %>% 
    filter(use_in_application == 1) %>% 
    mutate(start_year = str_extract(periode, "^\\d{4}")) %>% 
    select(register_name, table_name, variable_name, start_year)

random_registers <- read_csv(here("data/random-undocumented.csv"))

authorized_researchers <-
    read_csv(
        here("data/authorized-researchers.csv"),
        col_types = cols(.default = col_character())
    )

knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")
```

Registers requested are:

```{r dst-register-list, results = "asis"}
needed_registers %>% 
    filter(!is.na(register_id) & owner == "dst") %>% 
    mutate(across(where(is.character), ~if_else(is.na(.), "TO UPDATE", .))) %>% 
    glue_data("- {register_id}: {register_name_dk} / {register_name_english}") %>% 
    sort() %>% 
    cat(sep = "\n")
```

In the Appendix we listed the ATC codes from the prescription register and ICD-8
and -10 codes from the patient register that we need to classify diabetes 
and the diseases that may arise following a diabetes diagnosis.

<!-- Please also include information on external data or own survey data if such -->
<!-- data is part of the project. In such cases please include approval from the -->
<!-- Danish Protection Agency. -->

## Særligt vedr. lægemiddelsdata / Especially regarding drug data 

The medication data we are requested are used in order to identify cases of disease
(e.g. diabetes) that might not have been diagnosed and recorded in the patient register.
The list of requested ATC codes is found in the Appendix section.

<!--
Only to be filled in if the project is to use drug data. Research projects that
must use data from the Medicines Database must be approved by the Danish Health
and Medicines Authority. The approval process takes place via a contact person
in Forskningsservice. Drug data may only be used for the purpose described in
this project setting. If drug data is to be used for other purposes, a renewed
approval from the Danish Health and Medicines Authority is required.
-->

## Offentliggørelse af resultater / Publication of results

Results will be published in open access scientific outlets and presented at
scientific conferences.

## Perspektivering / Perspective

Understanding how early life and familial factors can influence diabetes provides
us with tools and knowledge to better target, intervene, and either
prevent or manage diabetes and its complications. Considering the impact that
diabetes as a disease itself and its related comorbidities has at the personal,
public health, and economic level, any improvement in knowledge can have major 
impacts.

<!-- A description of what knowledge of societal relevance this study can contribute. -->

## Data som ikke kommer fra Danmarks Statistik / Data that doesn't come from DST

Data coming from other sources (e.g. Sundhedsdatastyrelsen):

<!--
List cohorts:

- Diet, Cancer, and Health study. Next Generation Study
- ADDITION, ADDITION-PRO cohort studies
    - Some ideas: Family characteristics and diabetes control.
- Inter99
- CORE Trial (Check your health study. This is a household intervention)
- The Danish National Health survey (2010, 2013 and 2017)
    - See: https://www.danishhealthdata.com/find-health-data/Danskernes-sundhed Data
    administrator: National Institute of Public Health, SDU, application here:
    http://www.danskernessundhed.dk/Spoergeskema
-->
    
```{r non-dst-register-list, results = "asis"}
needed_registers %>% 
    filter(!is.na(register_id) & owner != "dst") %>% 
    mutate(across(where(is.character), ~if_else(is.na(.), "TO UPDATE", .))) %>% 
    glue_data("- {register_id}: {register_name_dk} / {register_name_english}") %>% 
    sort() %>% 
    cat(sep = "\n")
```

## Projekts slutdato / Project's end date

Date of deletion of data: Access to data is requested until 2030-12-31.

\newpage

## Autoriserede forskere / Authorised scientists

`r if (!knitr::is_html_output()) "
TODO: Add in relevant missing info before sending.
"`

```{r authorized-researchers-table}
library(flextable)
authorized_researchers_table <- function(data) {
    flextable(data) %>%
        border_inner() %>%
        border_outer() %>%
        flextable::bold(part = "header") %>%
        valign(valign = "top", part = "header") %>%
        bg(bg = "#d9d9d9", part = "header") %>%
        bg(i = c(1, 3), bg = "#f2f2f2") %>%
        flextable::bold(i = c(1, 3)) %>%
        merge_h_range(
            i = c(1, 3),
            j1 = 1,
            j2 = ncol(authorized_researchers)
        ) %>%
        flextable::fontsize(size = 9, part = "all") %>%
        {if (!knitr::is_html_output()) {font(., fontname = "Bitstream Vera Sans", part = "all")} else {.}}
}

authorized_researchers %>% 
    authorized_researchers_table()
```

\newpage

## Underskrift / Signatures

Underskriver bekræfter på vegne af den dataansvarlige autoriserede institution,
at den konkrete behandling er lovlig i henhold til databeskyttelsesforordningen
artikel 6.

Underskriver bekræfter ligeledes at være formelt ansat på den dataansvarlige
autoriserede institution, der indstiller projektet.

```{r}
signature_table <- function(name, position) {
    table_prep <- tribble(
        ~base, ~values,
        "Navn:", name,
        "Stilling:", position,
        "Dato:", "",
        "Underskrift:", ""
    )
    
    table_prep %>% 
        flextable::flextable() %>% 
        flextable::delete_part() %>% 
        flextable::border_remove() %>% 
        flextable::width(width = c(2, 4.5))
}
```


**På vegne af den dataansvarlige:**

`r if (!knitr::is_html_output()) "
TODO: Add signature before sending.
"`

```{r sign-table-1}
signature_table("Annelli Sandbæk", "Enhedschef")
```

**Godkendt til ekstern adgang**

```{r sign-table-2}
signature_table("Nikolaj Borg Burmeister", "Kontorchef")
```
