---
output: word_document
---

# Project description for DST application

## Autoriseret institution / Authorised institution

Steno Diabetes Center Aarhus (inst. nr. 574)  
Hedeager 3, 2. sal  
8200 Aarhus  
Denmark  

## Projekttitel / Project title

Interplay between diabetes and intergenerational transmission of health determinants over the life course

## Projektbeskrivelse / Project description

The overall aim of this project is to identify the contributions of family 
determinants on the development, management, and care of diabetes and the 
diseases that may arise following a diabetes diagnosis, under a life course
framework.
Diabetes is a complex, multi-factorial disease. Its development depends on 
the timing and accumulation of exposures throughout the life-course. In 
addition to the genetic influence, behavioral risk factors and socioeconomic
conditions cluster in families and are passed through generations. For instance,
poor childhood socioeconomic status (SES) can impact final educational attainment,
adult SES, and lifestyle and behavioral characteristics, which ultimately influences
an individual’s metabolic capacity to regulate glucose and subsequent risk for diabetes.
This project will contribute to a broader and deeper understanding of the role
that these factors play with the development and progression of diabetes. For 
this purpose, we will apply causal and predictive statistical methods to best
use the data available from the integrated data sources with an overarching goal 
to model several potential public health strategies that might impact the 
prevention and management of diabetes and the diseases that follows its diagnosis.
A key feature to this project is the integration of external data sources,
such as longitudinal cohort studies, with the register data.

## Population / Population

In order to identify familial relations through individuals’ life course trajectory,
the population requested covers all individuals, who live or lived in Denmark 
from 1 January 1970 to 31 December 2020 (~8.5 million people). Key requirements are:

- Individuals were born starting from 1970 with available parental linkage. Wherever
  possible, individuals and parental data throughout their early life, childhood 
  and adulthood.

- Minimum two and up to four-generation family pedigrees in which index individuals
  have linkage to their parents and spouses. Index individuals are defined as those 
  individuals that serve as a reference to identify family relations and which will
  be followed with regards to the outcomes.

## Variabelindhold / Variables included

All variables can be de-identified (direct identification is not needed). The
list of requested variables is found in the tables at the end of the document in 
the Appendix.
Registers requested are:
- AKM: Arbejdsklassifikationsmodulet / The work classification module
- BEF: Befolkningen / The population
- BYSTRB: Bystørrelser bopæle / City sizes residences
- DOD, DODSAASG and DODSAARS: Dødsårsagsregister / Cause of death register
- FAHU: Husstande / Households
- FAIK: Familieindkomster / Family income
- FAIN: Husstande og familier / Households and families
- FTBARN, FTDB, FTDK, FTDM, FTFORAEL, FTNAEVN: Fertilitetsdatabasen / Fertility database 
- IEPE: Indvandrere og Efterkommere / Immigrants and Descendants
- IND: Indkomst / Income
- LMDB: Lægemiddeldatabasen / The drug database
- LPR_ADM, LPR_BES, LPR_DIAG, LPR_OPR, LPR_SKSOP, LPR_SKSOPR, UAF_ADM, UAF_BES, UAF_DIAG,
  UAF_OPR, LPR_SKSUBE, UAF_SKSOPR: Landspatientregistret/ National Patient Register
- LPR_PSYK: Landspatientregisteret for psykiatri – centralpsykiatrisk register 
- LPRMFRDF: Dødfødte ud fra LPR / Stillbirths from LPR
- LPRMFRLF: Levendefødte ud fra LPR / Live births from LPR
- MFR: Levendefødte fra det medicinske fødselsregister / Live births from the 
  Medical Birth Register
- PRIV_ADM, PRIV_DIAG, PRIV_SKSOPR, PRIV_SKSUBE: Landspatientregistret private sygehuse
  / National Patient Register private hospitals 
- RAS: Registerbaserede arbejdsstyrkestatistik / Register-based labor force statistics
- SHSS: Sammenhængende socialstatistik / Coherent social statistics
- SSNV: Sammenhængende socialstatistik nettovarigheder / Coherent social statistics 
  net durations
- SSSY, SYSI, SYST: Sygesikringsregisteret / Helath Insurance Register
- UDDA: Uddannelser (BUE) / Education (BUE)
- VNDS: Historiske vandringer / Historical walks

Data som ikke kommer fra Danmarks Statistik / Data that doesn’t come from DST
List of studies:
- Diet, Cancer, and Health study. Next Generation Study
- ADDITION, ADDITION-PRO cohort studies
- Inter99
- CORE Trial (Check your health study. This is a household intervention)
- The Danish National Health survey (2010, 2013 and 2017)

Fra Sundhedsdatastyrelsen vil vi desuden søge om adgang til følgende register for data om diabetes samt relaterede sygdomme og risikofaktorer:
Kroniske sygdomsdiagnoser:
- Cancerregisteret
- IVF-registeret
- LABKA data
- Drug addicts in treatment and alcohol treatment registers
- Children data base

## Projekts slutdato / Project’s end date

Date of deletion of data: Access to data is requested until December 2031.

## Autoriserede forskere / Authorised scientists
*See the email to get the list


-->

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(glue)
# library(gt)
dst_variables <- read_csv(here("data/dst-registers-with-variables-to-use.csv"))
dst_variables <- dst_variables %>%
    filter(use_in_application == 1) %>% 
    mutate(across(ends_with("year"), str_extract, pattern = "^\\d{4}")) %>% 
    mutate(years = str_c(variable_start_year, "--", variable_end_year)) %>% 
    select(register_id, register_name_dk, variable_name, years)

authorized_researchers <-
    read_csv(
        here("data/authorized-researchers.csv"),
        col_types = cols(.default = col_character())
    )

knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")
```

Registers requested are:

```{r register-list, results = "asis"}
read_csv(here("data/needed-registers.csv"), col_types = "dccccc") %>% 
    filter(!is.na(register_id)) %>% 
    mutate(across(where(is.character), ~if_else(is.na(.), "TO UPDATE", .))) %>% 
    glue_data("- {register_id}: {register_name_dk} / {register_name_english}") %>% 
    sort() %>% 
    cat(sep = "\n")
```

<!-- Please also include information on external data or own survey data if such -->
<!-- data is part of the project. In such cases please include approval from the -->
<!-- Danish Protection Agency. -->

## Særligt vedr. lægemiddelsdata / Especially regarding drug data 

<!--
Only to be filled in if the project is to use drug data. Research projects that
must use data from the Medicines Database must be approved by the Danish Health
and Medicines Authority. The approval process takes place via a contact person
in Forskningsservice. Drug data may only be used for the purpose described in
this project setting. If drug data is to be used for other purposes, a renewed
approval from the Danish Health and Medicines Authority is required.
-->

## Offentliggørelse af resultater / Publication of results

Results will be published in open access scientific outlets and presented at
scientific conferences.

## Perspektivering / Perspective

<!-- A description of what knowledge of societal relevance this study can contribute. -->

## Data som ikke kommer fra Danmarks Statistik / Data that doesn't come from DST

List cohorts:

- Diet, Cancer, and Health study. Next Generation Study
- ADDITION, ADDITION-PRO cohort studies
    - Some ideas: Family characteristics and diabetes control.
- Inter99
- CORE Trial (Check your health study. This is a household intervention)
- The Danish National Health survey (2010, 2013 and 2017)
    - See: https://www.danishhealthdata.com/find-health-data/Danskernes-sundhed Data
    administrator: National Institute of Public Health, SDU, application here:
    http://www.danskernessundhed.dk/Spoergeskema
    
Fra Sundhedsdatastyrelsen vil vi desuden søge om adgang til følgende register
for data om diabetes samt relaterede sygdomme og risikofaktorer:

- Kroniske sygdomsdiagnoser:
- Det Nationale Diabetesregister? (LPR, RMPS, DADD, Diabase)
- Medicinering kronisk sygdom:
- Exposures and risk factors: Overvægt?
- LPR diagnosis of overweight may be included : ICD 10: E66. Maybe possible to
get some data from electronic health records.
- drug addicts in treatment and alcohol treatment registers. Could we use
diagnoses of substance abuse / alcohol abuse: ICD-10: F1x
- children data base

## Projekts slutdato / Project's end date

Date of deletion of data: Access to data is requested until December 2030.

## Autoriserede forskere / Authorised scientists

```{r}
knitr::kable(authorized_researchers)
```

# Bilag / Appendix

## List of registers and variables, with explanations

```{r requested-variable-table, echo=FALSE, results='asis'}
variable_tables<- dst_variables %>% 
    mutate(register_name = glue("{register_name_dk} ({register_id})")) %>% 
    select(-register_id, -register_name_dk) %>% 
    rename(Variables = variable_name, Years = years) %>% 
    group_split(register_name) 

for (variable_table in variable_tables) {
    variable_table %>% 
        select(-register_name) %>%  
        knitr::kable(caption = str_c(
            "Requested variables from ",
            unique(variable_table$register_name)
        )) %>% 
        print()
}
    # gt() %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_row_groups()
    # ) %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_column_labels(everything())
    # ) %>% 
    # tab_header("Explanation for requested selected variables and registers.") %>% 
    # cols_label(variable_english = "Variable in English", explanation = "Explanation")
```
