---
output: word_document
---

# Project description for DST application

## Autoriseret institution / Authorised institution

Steno Diabetes Center Aarhus  
Hedeager 3, 2. sal  
8200 Aarhus  
Denmark  

## Projekttitel / Project title

MELD project: Metabolic consequences of early life adversity and subsequent risk for type 2 diabetes

## Projektbeskrivelse / Project description

Early life adversity is known to influence the risk for adult cardiometabolic
diseases such as type 2 diabetes (T2D); however, how metabolic processes mediate
this link is not well understood. Much of what we know about how early life
metabolically influences later risk for chronic disease comes from non-human 
animal research. With the availability of long-term longitudinal data from the
registers, linked to cohort datasets with metabolic data, we aim to:

1. Investigate how and to what degree individual early life conditions may
contribute to the risk of T2D in adulthood.
1. Examine potential mediating pathways that adult metabolic capacity may
have between exposure to adversity in early life and risk for incident T2D,
along with the estimated magnitude of the pathways.

See Appendix 1 for a more complete project description (in English).

TODO: Revise DDA project description and add as appendix item

## Population / Population

The exposure time period of interest is from conception until about 7 years of age.
So the population of interest would be all individuals born after (1968?) until
the present time that have data on their parents 1 year before the
individual's birth and until 7 years of age.

TODO: Justify why here. Maybe give an estimate to population number?

## Variabelindhold / Variables included

```{r setup, include=FALSE}
library(tidyverse)
# library(gt)
variable_list <- read_csv(here::here("data/registries-general-variables.csv"))
authorized_researchers <-
    read_csv(
        here::here("data/authorized-researchers.csv"),
        col_types = cols(.default = col_character())
    )
knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")
```

Registers requested are:

Personoplysninger

- BEF
- FAIN
- VNDS
- FAIK
- IND
- AKM
- UDDA
- FAM
- HUST
- FAHU
- IEPE
- VNDS
- IDA
- RAS
- PUOB 
- INDH
- INDK
- IEPE

Sygesikring

- SYSI
- SSSY
- SYST
- SHSS
- SSNV

Landspatientregistret indlæggelser (administrative oplysninger)

- LPR_ADM

LPRBES (ambulant besøg)

- LPR_BES

LPRDIAG (diagnose)

- LPR_DIAG

Operation

- LPR_OPR
- LPR_SKSOP
- LPR_SKSUBE

Privathospitals indlæggelser (administrative oplysninger)

- PRIV_ADM

PRIVDIAG (Privathospital diagnose)

- PRIV_DIAG

PRIVSKSUBE (Privathospital undersøgelser og behandlinger)

- PRIV_SKSUBE

PRIVSKSOPR (Privathospital operation)

- PRIV_SKSOPR

Landspatientregistret uafsluttet indlæggelser (administrative oplysninger)

- UAF_ADM
- UAF_BES
- UAF_DIAG

Uafsluttet operation

- UAF_OPR
- UAF_SKSOPR

Døde i Danmark

- DOD
- DODSAASG

Fra Sundhedsdatastyrelsen

Lægemiddelstatistikregisteret (LMSREG)

- LMDB

Landspatientregistret-Psykiatri

- LPR_PSYK

LABKA

Geografi, miljø og energi:

- BYSTRB



```{r register-list, results="asis"}
variable_list %>% 
    glue::glue_data("- {register_english} / {register_danish} ({register_abbreviation})") %>% 
    unique() %>%
    str_remove(" \\(NA\\)") %>% 
    str_remove(" \\/ NA") %>% 
    sort() %>% 
    cat(sep = "\n")
```

General variables requested are:

- Specific exposure variables during individual's exposure time period: 
    - Parental socio-economic
        - Household income
        - Educational attainment
        - Parental unemployment
        - Household poverty
        - Social housing
    - Parental health:
        - Death
        - Severe illness (how to define that further?)
        - Severe psychiatric illness
        - Alcohol and drug abuse
    - Maternal:
        - Smoking during pregnancy and breastfeeding
        - Breastfeeding
    - Parental conditions
        - Divorce
        - Separation
        - Single parent
        - Spousal abuse or injuries
        - Number of times moved
    - Emigration
    - Neighbourhood-level:
        - General income/poverty
        - General crime/safety (possible?)
    - Individual:
        - Severe illness or injury
        - Foster care
    - Sibling:
        - Psychiatric illness
        - Severe illness

- Potential exposure indicators:
    - Birthweight
    - Birth length
    - Height from multiple periods in life
    - Leg length (from cohorts)
    - Weight from multiple periods in life

- Other variables:
    - Year of birth
    - Month of birth
    - Birth order
    - Number of siblings
    - Parental age
    - Ethnicity
    - Sex
    - Medical data from pregnancy, at birth, and in first years
    - Date of birth

- Outcome variable:
    - Age at diabetes diagnosis 
    - Diabetes status at any given time

Exact variables requested from each register are listed in the section below
along with the justification for requesting them.

<!-- Please also include information on external data or own survey data if such -->
<!-- data is part of the project. In such cases please include approval from the -->
<!-- Danish Protection Agency. -->

## Data som ikke kommer fra Danmarks Statistik / Data that doesn't come from DST

List cohorts:

## Projekts slutdato / Project's end date

Date of deletion of data: Access to data is requested until December 2023.

## Autoriserede forskere / Authorised scientists

```{r}
knitr::kable(authorized_researchers)
```

# List of registers and variables, with explanations

```{r justification-table, echo=FALSE}
variable_list %>% 
    dplyr::filter(use_in_study == "yes") %>% 
    group_by(register_english) %>% 
    mutate(Register = ifelse(row_number() == 1, register_english, "")) %>% 
    ungroup() %>% 
    select(Register, Variable = variable_english, Explanation = explanation) %>% 
    knitr::kable(caption = "Explanation for requested selected variables and registers.")
    # gt() %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_row_groups()
    # ) %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_column_labels(everything())
    # ) %>% 
    # tab_header("Explanation for requested selected variables and registers.") %>% 
    # cols_label(variable_english = "Variable in English", explanation = "Explanation")
```
