---
output: odt_document
---

# Project description for DST application

## Autoriseret institution / Authorised institution

Steno Diabetes Center Aarhus  
Hedeager 3, 2. sal  
8200 Aarhus  
Denmark  

## Projekttitel / Project title

MELD project: Metabolic consequences of early life adversity and subsequent risk for type 2 diabetes

## Projektbeskrivelse / Project description

Early life adversity is known to influence the risk for adult cardiometabolic
diseases such as type 2 diabetes (T2D); however, how metabolic processes mediate
this link is not well understood. Much of what we know about how early life
influences later risk for chronic disease comes from

With the availability of long-term longitudinal data from the registers,Given the 
metabolism and to T2D risk. My aims are to 1) understand how specific early life
conditions affect T2D risk, 2) identify the metabolic pathways that mediate
between early life adversity and adult T2D risk, and 3) to develop a recent
causal pathway algorithm (NetCoupler) capable of performing mediation and causal
structure analysis on detailed metabolic data into a reusable
software package. To achieve these aims, I will a) use Denmark's unique national
registers linked with multiple cohorts, b) deploy and apply NetCoupler, and c)
collaborate with national and international groups. The software will provide a
tool for researchers working in similar areas, while the results will give
insight into how more effective public health strategies could be created for
T2D prevention and how more precise clinical decisions could be made for T2D
management in those exposed to early life adversity.

**Objectives:**

1. To investigate how individual early life conditions affect the risk of
T2D in adulthood.
1. To rank the contributions that various early life conditions have on risk
for T2D.
1. To examine any potential mediating pathways that adult metabolic capacity may
have between exposure to adversity in early life and risk for incident T2D.
1. To estimate the magnitude of these mediating pathways and rank the contributions
these pathways have between early life adversity and risk for T2D.

See Appendix 1 for a more complete project description (in English).

## Population / Population

Anyone born after (1968?) until the present time as well as information on their
parents 1 year before an individual's birth and until 7 years of age. Since we
are t

(Justify why here. Maybe give an estimate to population number?)

## Variabelindhold / Variables included

```{r setup, include=FALSE}
library(tidyverse)
library(gt)
variable_list <- read_csv(here::here("data/registries-general-variables.csv"))
authorized_researchers <-
    read_csv(
        here::here("data/authorized-researchers.csv"),
        col_types = cols(.default = col_character())
    )
knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")
```

Registers requested are:

```{r register-list, results="asis"}
variable_list %>% 
    glue::glue_data("- {register_english} / {register_danish} ({register_abbreviation})") %>% 
    unique() %>%
    str_remove(" \\(NA\\)") %>% 
    sort() %>% 
    cat(sep = "\n")
```

Exact variables requested from each register are listed in the section below
along with the justification for requesting them.

<!-- Please also include information on external data or own survey data if such -->
<!-- data is part of the project. In such cases please include approval from the -->
<!-- Danish Protection Agency. -->

## Data som ikke kommer fra Danmarks Statistik / Data that doesn't come from DST

List cohorts:

## Projekts slutdato / Project's end date

Date of deletion of data: Access to data is requested until December 2023.

## Autoriserede forskere / Authorised scientists

```{r}
knitr::kable(authorized_researchers)
```

# List of registers and variables, with explanations

```{r justification-table, echo=FALSE}
variable_list %>% 
    dplyr::filter(use_in_study == "yes") %>% 
    group_by(register_english) %>% 
    mutate(Register = ifelse(row_number() == 1, register_english, "")) %>% 
    ungroup() %>% 
    select(Register, Variable = variable_english, Explanation = explanation) %>% 
    knitr::kable(caption = "Explanation for requested selected variables and registers.")
    # gt() %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_row_groups()
    # ) %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_column_labels(everything())
    # ) %>% 
    # tab_header("Explanation for requested selected variables and registers.") %>% 
    # cols_label(variable_english = "Variable in English", explanation = "Explanation")
```
