---
output: word_document
---

# Project description for DST application

## Autoriseret institution / Authorised institution

Steno Diabetes Center Aarhus (inst. nr. 574)  
Hedeager 3, 2. sal  
8200 Aarhus  
Denmark  

## Projekttitel / Project title

Family relations and type 2 diabetes through the life span

## Projektbeskrivelse / Project description

The overall aim of this project is to identify the contribution of family
characteristics in terms of chronic disease history, socio demographic
characteristics and risk factors for the development of type 2 diabetes under a
life course perspective using a three-generation family pedigree.

Type 2 diabetes is a multifactorial disease that can develop at any stage of
life course trajectory; yet it is most commonly diagnosed during middle-age
adulthood. In addition to the strong genetic risk component, behavioral risk
factors and socio economic conditions clusters in families and their influence
depends on the timing and exposure accumulation (length). For example: poor
childhood socio economic status (SES) --> poor educational attainment --> poor
adult SES poor childhood SES --> poor quality diet --> lower pancreatic
potential **** add a simple DAG of your project instead. Use these three nodes
structure to keep it simple***

This project will contribute with new and strong evidence on the role that
family relations plays for type 2 diabetes development and progression under an
aetiological, predictive and a causal perspective using advanced
epidemiological, demographic, and statistical methods.


## Population / Population

In order to identify familial relations through individuals’ life course
trajectory, the population understudy covers all individuals alive and living in
Denmark from 1 January 1970 to 31 December 2020 (~8.5 million people). It
consist of a two or three-generation (even four) pedigree in which index
individuals have linkage to their parents and spouses. Index individuals are
defined as those individuals that serve as a reference to identify family
relations and which will follow with regards to the outcome. With the purpose to
identify potential pathways, drivers and mechanisms linking family relations to
type 2 diabetes development and progression, this project will further use
cohort and cross-sectional survey data to extract more detailed information on
individuals’ and family members risk exposures.

## Variabelindhold / Variables included

All variables can be de-identified (direct identification is not needed).

- Befolkning vs. Demografi, familie og husstandsforhold (BEF) 
- BEF_ADRESSELABEL, BEF_PERSONNUMMER, Byopgørelsen, Familier og husstande
- Fertilitetsdatabasen
- Personindkomst vs. Indkomster (INDH/INDK)
- Familieindskomst (FAIK)
- Kvalifikationsregistret vs. Uddannelsesoplysninger (UDDA) 
- Vandringer (VNDS)
- Arbejdsklassifikationsmodulet (AKM)
- Sygesikringsregisteret (SYSI, SSSY)
- Lægemiddelregisteret (LMDB)
- Arbejdsplads(er)

##Just a random though, early life stressor: parents' imprisonment. Perhaps you want to have a look 
at this registers: 
Børnetilskud og børnefamilieydelse
Levevilkårsundersøgelsen (SILC)
Forbrugsundersøgelsen?
Idræt?

## Særligt vedr. Lægemiddelsdata
Der ønskes adgang til lægemidler, da oplysninger om medicin bidrager til at identificere personer med
diabetes og derudover bidrager til at belyse sygelighed i populationen. 

Eksterne data eller egne data
Eksterne data/egne data til dette projekt er:
Fra Sundhedsdatastyrelsen vil vi desuden søge om adgang til følgende register for data om diabetes
samt relaterede sygdomme og risikofaktorer:

- Kroniske sygdomsdiagnoser:
Det Nationale Diabetesregister? (LPR, RMPS, DADD, Diabase)
- Landspatientregistret 
Cancerregisteret
The laboratory data set/ Danish National Biobank?
- Dødsårsag:
Dødsårsagsregisteret (DOD, DODSAARS, DODSAASG)
- Medicinering kronisk sygdom:
Lægemiddelstatistikdatabasen
- Exposures and risk factors:
Overvægt:Sessionsregisteret?

```{r setup, include=FALSE}
library(tidyverse)
# library(gt)
variable_list <- read_csv(here::here("data/registries-general-variables.csv"))
authorized_researchers <-
    read_csv(
        here::here("data/authorized-researchers.csv"),
        col_types = cols(.default = col_character())
    )
knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")
```

Registers requested are:

```{r register-list, results="asis"}
variable_list %>% 
    glue::glue_data("- {register_english} / {register_danish} ({register_abbreviation})") %>% 
    unique() %>%
    str_remove(" \\(NA\\)") %>% 
    str_remove(" \\/ NA") %>% 
    sort() %>% 
    cat(sep = "\n")
```

<!-- Please also include information on external data or own survey data if such -->
<!-- data is part of the project. In such cases please include approval from the -->
<!-- Danish Protection Agency. -->

## Data som ikke kommer fra Danmarks Statistik / Data that doesn't come from DST

List cohorts:

- EPIC-Interact cohort study. Next Generation Study
- ADDITION, ADDITION-PRO cohort studies
- Inter99
- CORE Trial (Check your health study)
- The Danish National Health survey (2010, 2013 and 2017)

## Projekts slutdato / Project's end date

Date of deletion of data: Access to data is requested until December 2030.

## Autoriserede forskere / Authorised scientists

```{r}
knitr::kable(authorized_researchers)
```

# List of registers and variables, with explanations

```{r justification-table, echo=FALSE}
variable_list %>% 
    dplyr::filter(use_in_study == "yes") %>% 
    group_by(register_english) %>% 
    mutate(Register = ifelse(row_number() == 1, register_english, "")) %>% 
    ungroup() %>% 
    select(Register, Variable = variable_english, Explanation = explanation) %>% 
    knitr::kable(caption = "Explanation for requested selected variables and registers.")
    # gt() %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_row_groups()
    # ) %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_column_labels(everything())
    # ) %>% 
    # tab_header("Explanation for requested selected variables and registers.") %>% 
    # cols_label(variable_english = "Variable in English", explanation = "Explanation")
```
