---
output: word_document
---

# Project description for DST application

## Autoriseret institution / Authorised institution

Steno Diabetes Center Aarhus  
Hedeager 3, 2. sal  
8200 Aarhus  
Denmark  

## Projekttitel / Project title

MELD project: Metabolic consequences of early life adversity and subsequent risk for type 2 diabetes

## Projektbeskrivelse / Project description

Early life adversity is known to influence the risk for adult cardiometabolic
diseases such as type 2 diabetes (T2D); however, how metabolic processes mediate
this link is not well understood. Much of what we know about how early life
metabolically influences later risk for chronic disease comes from non-human 
animal research. With the availability of long-term longitudinal data from the
registers, linked to cohort datasets with metabolic data, we aim to:

1. Investigate how and to what degree individual early life conditions may
contribute to the risk of T2D in adulthood.
1. Examine potential mediating pathways that adult metabolic capacity may
have between exposure to adversity in early life and risk for incident T2D,
along with the estimated magnitude of the pathways.

See Appendix 1 for a more complete project description (in English).

TODO: Revise DDA project description and add as appendix item

## Population / Population

The exposure time period of interest is from conception until about 7 years of age.
So the population of interest would be all individuals born after (1968?) until
the present time that have data on their parents 1 year before the
individual's birth and until 7 years of age.

TODO: Justify why here. Maybe give an estimate to population number?

## Variabelindhold / Variables included

```{r setup, include=FALSE}
library(tidyverse)
# library(gt)
variable_list <- read_csv(here::here("data/registries-general-variables.csv"))
authorized_researchers <-
    read_csv(
        here::here("data/authorized-researchers.csv"),
        col_types = cols(.default = col_character())
    )
knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")
```

Registers requested are:

```{r register-list, results="asis"}
variable_list %>% 
    glue::glue_data("- {register_english} / {register_danish} ({register_abbreviation})") %>% 
    unique() %>%
    str_remove(" \\(NA\\)") %>% 
    str_remove(" \\/ NA") %>% 
    sort() %>% 
    cat(sep = "\n")
```

<!-- Please also include information on external data or own survey data if such -->
<!-- data is part of the project. In such cases please include approval from the -->
<!-- Danish Protection Agency. -->

## Data som ikke kommer fra Danmarks Statistik / Data that doesn't come from DST

List cohorts:

## Projekts slutdato / Project's end date

Date of deletion of data: Access to data is requested until December 2023.

## Autoriserede forskere / Authorised scientists

```{r}
knitr::kable(authorized_researchers)
```

# List of registers and variables, with explanations

```{r justification-table, echo=FALSE}
variable_list %>% 
    dplyr::filter(use_in_study == "yes") %>% 
    group_by(register_english) %>% 
    mutate(Register = ifelse(row_number() == 1, register_english, "")) %>% 
    ungroup() %>% 
    select(Register, Variable = variable_english, Explanation = explanation) %>% 
    knitr::kable(caption = "Explanation for requested selected variables and registers.")
    # gt() %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_row_groups()
    # ) %>% 
    # tab_style(
    #     style = cell_text(weight = "bold"),
    #     locations = cells_column_labels(everything())
    # ) %>% 
    # tab_header("Explanation for requested selected variables and registers.") %>% 
    # cols_label(variable_english = "Variable in English", explanation = "Explanation")
```
