---
title: "Application for use of Sundhedsdatastyrelsen (SDS) registers"
author:
  - name: Luke W. Johnston
    url: https://lukewjohnston.com/
    affiliation: Steno Diabetes Center Aarhus
repository_url: https://gitlab.com/lwjohnst/meld-dst-application/
creative_commons: CC BY
---

Fill in the below information using the [SDS application
form](https://ds.sst.dk/ds2010/Forskerservice_CRM2/Forskerservice.aspx).
Not all information filled into the form is present here.

-   **Project number**: 708421

# Project title

Interplay between diabetes and intergenerational transmission of health
determinants over the life course

(Max 200 characters)

# Abstract

Diabetes is a complex, multi-factorial disease. Its development depends
on the timing and accumulation of exposures throughout the life-course.
In addition to genetic influences, behavioral risk factors and
socioeconomic conditions cluster in families and are passed through
generations. The overall aim of this project is to identify the
contributions of family and early life determinants on the development,
management, economics, and care of diabetes and the diseases that may
arise following a diabetes diagnosis, under a life-course framework.

# Registers

```{r import-registers-needed, include=FALSE}
library(tidyverse)
library(here)
library(glue)
sds_variables <- read_csv(here("data/sds-registers-with-variables-to-use.csv"))
sds_variables <- sds_variables %>% 
    filter(use_in_application == 1) %>% 
    mutate(start_year = as.numeric(str_extract(periode, "^\\d{4}"))) %>% 
    select(register_name, table_name, variable_name, start_year)

random_registers <- read_csv(here("data/random-undocumented.csv"))
```

```{r list-registers-needed, echo=FALSE, results='asis'}
bind_rows(sds_variables,
          random_registers) %>%
    distinct(register_name) %>% 
    filter(register_name %in% c("Cancerregisteret", "IVF", "LPR_PSYK", "DRG", "PLH", "RUKS")) %>% 
    add_row(register_name = "LMDB") %>% 
    arrange(register_name) %>% 
    glue::glue_data("- {register_name}")
```
