---
title: "Requested ATC codes"
---

## Justifications

-   Many of the detailed level 5 ATC codes are used to determine
    diabetes or cardiovascular disease status, or to help identify
    co-morbidities from these diseases.
-   Many of the level 2 ATC codes are needed to calculate the total cost
    to the individual. This is standard to include all prescriptions and
    not only the ones related to diabetes. We use the level 2 ATC codes
    to make a rough classification of the groups of medicine used.

{{< pagebreak >}}

```{r list-atc-codes, include=FALSE}
library(here)
library(tidyverse)
library(flextable)
source(here("R/format-tables.R"))

options(knitr.kable.NA = "")
# To get loop to work
if (knitr::is_html_output())
    htmltools::div(flextable_html_dependency())

atc_codes <- here("data/atc-codes.csv") %>%
    read_csv(col_types = cols(
        level = col_double(),
        code = col_character(),
        description = col_character(),
        expanded_level = col_double()
    )) %>% 
    mutate(
        level = if_else(
            is.na(expanded_level),
            level,
            expanded_level
        ),
    ) %>% 
    # janitor::get_dupes(level, code)
    select(level, code) 

atc_codes_list <- atc_codes %>% 
    arrange(level) %>% 
    group_split(level) %>% 
    map(
        ~ .x %>%
            mutate(
                `LMDB variable` = str_replace(glue::glue("ATC{level}"), "ATC5", "ATC"),
                `LMDB variable` = if_else(row_number() == 1, `LMDB variable`, ""),
                renamed_column = glue::glue("ATC-Code ({level}. niveau)")
            ) %>%
            pivot_wider(
                names_from = renamed_column,
                values_from = code,
                values_fn = list
            ) %>%
            unnest(cols = everything()) %>% 
            select(-level) 
        )
```

```{r table-atc-codes}
#| results: asis
#| tbl-cap: "ATC Codes."
# Need a loop for working with flextables
for (atc_code_table in atc_codes_list) {
    cat("\n\n")
    atc_code_table %>% 
        format_non_dst_appendix_table(width = c(2, 5))
}
```
