---
title: "Data extraction description"
author:
  - name: Luke W. Johnston
    url: https://lukewjohnston.com/
    affiliation: Steno Diabetes Center Aarhus
repository_url: https://gitlab.com/lwjohnst/meld-dst-application/
creative_commons: CC BY
output: 
    word_document:
        reference_docx: ../../resources/templates/appendix.docx
---

```{=html}
<!-- 
Your extract description must consist of the following points: 

- Place of treatment
- Population
- Data connection
- Justification for the requested data
-->
```
## Data access and user on the Research Machine

We do not need access to a research machine, we have a project at DST.

## Treatment site

We want the data to be transferred to our project on the DST research
server.

## Population

<!-- State whether you need CPR linked for you or if you will be doing it. -->

To identify familial relations through individuals' life course
trajectory, the population requested covers all individuals, who live or
lived in Denmark from 1 January 1970 to 31 December 2021 (\~8.5 million
people). The family linkage and structure makes the construction of the
population complicated, which we will do ourselves. For that we need the
entire population because we need data on not just the individual, but
also the family (parents and siblings). Denmark Statistics has given us
the population and data we requested.

## Data connection

```{=html}
<!-- 
You must specify what data you want to access for the population you have specified. 

Data must as a minimum be delimited on the basis of the following parameters:

- Which registers
- Which tables
- Which variables
- What time period
-->
```
```{r list-registers-needed, include=FALSE}
library(tidyverse)
library(here)
needed_registers <- here("data/needed-registers.csv") %>% 
    read_csv(col_types = "lddccccc") %>% 
    filter(!drop | is.na(drop))

dst_register_names <- c("Lægemiddeldatabasen") 
dst_variables <- read_csv(here("data/dst-registers-with-variables-to-use.csv"))
dst_variables <- dst_variables %>%
    filter(use_in_application == 1,
           register_id %in% unique(needed_registers$register_id),
           register_name_dk %in% dst_register_names) %>% 
    mutate(across(ends_with("year"), str_extract, pattern = "^\\d{4}")) %>% 
    mutate(
        start_year = case_when(
            is.na(variable_start_year) |
                variable_start_year == variable_end_year ~
                register_start_year,
            TRUE ~ variable_start_year
        ),
        time_period = glue::glue("{start_year} to 2021")
    ) %>% 
    select(register_name = register_name_dk, variable_name, time_period)

sds_variables <-
    read_csv(here("data/sds-registers-with-variables-to-use.csv"),
             show_col_types = FALSE) %>% 
    filter(use_in_application == 1) %>% 
    mutate(
        periode = periode %>%
            str_trim() %>%
            str_replace_all(" +", " ") %>%
            str_replace_all(" ?- ?", " to ") %>%
            str_replace(" og ", ", "),
        time_period = case_when(
            str_detect(periode, "to $") ~ glue::glue("{str_remove(periode, ' -$')} 2021"),
            TRUE ~ periode
        )
    ) %>% 
    select(register_name, table_name, variable_name, time_period)

sds_register_names <- c("Cancerregisteret", "IVF", "LPR_PSYK", "DRG", "PLH", "RUKS", "LAB")
registers <- bind_rows(sds_variables, random_registers) %>% 
    filter(register_name %in% sds_register_names) %>% 
    bind_rows(dst_variables)
```

```{r tables, echo=FALSE, results='asis'}
header_and_table_as_md <- function(data) {
    header <- unique(data$register_name)
    
    table <- data %>% 
        select(-register_name) %>% 
        select(where(~all(!is.na(.x)))) %>% 
        rename_with(~str_replace_all(.x, "_", " ") %>% str_to_sentence()) %>% 
        knitr::kable() %>% 
        as.character()
    
    c("", glue::glue("## {header}"), "", table, "")
}

registers %>% 
    group_split(register_name) %>% 
    map(header_and_table_as_md) %>% 
    unlist() %>% 
    cat(sep = "\n")
```

## Justification for the requested data

```{=html}
<!-- 
We ask you to briefly describe how the desired data can be used to answer the
project's problem.

This is to ensure that you as a researcher have access to data that is both
necessary and sufficient for your project.
-->
```
-   Population:

    -   We require data on all individuals, including those without
        diabetes, in order for our models to have a comparison to use
        and to identify risk factors. We also need the data on the
        individuals *before* they get diabetes until their birth in
        order to study life course effects on diabetes risk and eventual
        prevention.

-   IVF treatments and cancer diagnoses on all persons who have lived
    and lived in Denmark in the period 1970-2021, including persons who
    are not diabetic patients or in a family relationship to a diabetic
    patient:

    -   Over the last decades, increasing trends in the incidence of
        specific types of cancers, obesity, and type 2 diabetes has been
        observed globally. Lifestyle factors such as the type of diet
        that is consumed, and the quantity of physical activity that is
        performed are some of the drivers in the development of obesity,
        which is a strong independent risk factor for type 2 diabetes
        and the link to certain types of cancers such as liver,
        pancreas, breast, among others. Furthermore, shared biological
        pathways between diabetes and cancer include inflammation,
        metabolic and genetic anomalies. In summary, a broad body of
        literature shows a strong link between diabetes and cancer. In
        this project, we would investigate how diabetes and cancer (and
        other diseases) cluster in families and how this disease
        aggregation in the family affects the individual risk of
        developing different chronic disease profiles (e.g. diabetes or
        cancer, or both). In short, data on cancer diagnoses on all
        persons who have lived and lived in Denmark from the period
        1970-2021 is necessary, as this population will serve as the
        reference group when relevant.

        For example, if we want to estimate the risk of diabetes for
        50-year-old male individuals with a paternal history of cancer
        and diabetes, the reference group would be the 50-year-old male
        individuals with no paternal history of cancer and diabetes. It
        would be possible to investigate the independent effect of
        paternal cancer history on diabetes risk among those individuals
        with no paternal history of diabetes. A second example of a
        reference population is when estimating mortality rates of
        diabetes and cancer using transition models. Where individuals
        enter the cohort without any disease and then move into the next
        state that could be either diabetes only, cancer only, or death,
        then move to diabetes with cancer and then transition to death.

-   Both CAR and IVF:

    -   Diseases following the diabetes diagnosis includes not only the
        classical diabetes related diseases, but also less studied
        diseases like reproductive impairment and cancers. Therefore, we
        request access to several registers to identify relevant factors
        related to family and early life determinants on the
        development, management, and care of diabetes and for
        development of diabetes and diseases following diabetes.

-   DRG og DAGS grupperet LPR (DRG)

    -   Hospital treatment is a crucial cost when it comes to diabetes
        care. It is important to analyse all costs (and not only those
        with an acute diagnosis of diabetes) as diabetes may be related
        to a general change in and therefore the consumption of hospital
        care.

    -   The years are extended to 2004 in order to calculate charlson
        co-morbidity index with a 10-year look back.

-   Landspatientregisteret psykiatri (LPR)

    -   Treatment in a psychiatric hospital may be influenced by a
        possible diabetes diagnosis and should therefore be included to
        provide the full picture.

-   Plejehjemsadresser (PLH)

    -   People with diabetes may be at increased risk of needing to live
        in a care home, particularly in the context of diabetes-related
        conditions. We need nursing home names in order to make the most
        detailed calculation of the cost of living in nursing home.

-   Register for Udvalgte Kroniske Sygdomme (RUKS)

    -   RUKS is used to compare the identified diabetes population and
        to compare morbidity between the identified diabetes population
        and the diabetes-free population.

-   Lægemiddelstatistiskregisteret hos Danmarks Statistisk (LMDB)

    -   Diabetes patients can have a considerable consumption of
        medicines both directly related to diabetes, but also other
        medicines.

-   Laboratoriedatabasen (LAB)

    -   HbA1c measurements from the laboratory database researcher table
        are desired to examine incidence and progression of diabetes in
        the whole population, especially from general practice which is
        only available to a limited extent. Data unbiased by value are
        desired as diagnostic criteria have changed over time. The
        remaining measurements (cholesterol, eGFR, CRP, Urat, etc.) are
        desired to examine the incidence and progression of
        complications of the disease. Information on the laboratory and
        the applicant is requested in order to take into account
        possible differences in measurements and results among
        laboratories and geographically. Reference ranges are requested
        to examine responses indicated to be in the normal range and any
        differences between laboratories and over time.
    -   The relevant NPU/DNK codes are listed in the table below.

## Requested LAB NPU/DNK codes

```{r labka-data-source, results='asis', echo=FALSE}
# Labka is same as LAB
here("data/labka.csv") %>%
    read_csv(col_types = cols(
        Komponent = col_character(),
        Kode = col_character(),
        Enhed = col_character()
    )) %>% 
    group_by(Komponent) %>% 
    mutate(Komponent = if_else(row_number() == 1, Komponent, "")) %>% 
    ungroup() %>% 
    knitr::kable()
```
